---
title: "AoC 2015 Day 1"
output: github_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
#knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
```

## PART 1
Santa is trying to deliver presents in a large apartment building, but he can't find the right floor - the directions he got are a little confusing. He starts on the ground floor (floor 0) and then follows the instructions one character at a time.

An opening parenthesis, (, means he should go up one floor, and a closing parenthesis, ), means he should go down one floor.

The apartment building is very tall, and the basement is very deep; he will never find the top or bottom floors.

For example:

(()) and ()() both result in floor 0.
((( and (()(()( both result in floor 3.
))((((( also results in floor 3.
()) and ))( both result in floor -1 (the first basement level).
))) and )())()) both result in floor -3.
To what floor do the instructions take Santa?

------------------
This is the input I was given:
```{r, warning = FALSE, echo=FALSE}
input <- read.table("Day1_Input1.csv")
input <- input$V1
print(input)
```

A quick and dirty solution would be to count the nos of '(' and ')' and then take the difference to determine the floor.

```{r }
up   <- str_count(input, "\\(")
down <- str_count(input, "\\)")
floor_final <- up - down 
```

Santa goes `up = `r up`` floors and `down = `r down`` floors taking him to `floor_final = `r floor_final``.

## PART 2
Now, given the same instructions, find the position of the first character that causes him to enter the basement (floor -1). The first character in the instructions has position 1, the second character has position 2, and so on.

For example:

) causes him to enter the basement at character position 1.
()()) causes him to enter the basement at character position 5.
What is the position of the character that causes Santa to first enter the basement?

```{r}
input_long <- str_split(input, "")[[1]]
# Loop through the input
input_len <- length(input_long)
floor <- vector(mode = "integer", length=input_len+1)
floor[1] <- 0
for(i in 1:input_len){
  if(input_long[i] == "("){
    floor[i+1] <- floor[i] + 1
    }
  else{
    if(input_long[i] == ")"){
      floor[i+1] <- floor[i] - 1
    }
    else print("error - not '(' or ')'")
  }
}
```



On which moves/characters did Santa reach the basement? Need to remember to subtract 1 since we added an initial entry to the `floor` vector so that it started from floor 0 
```{r}
which(floor==-1) - 1
```

Plot Santa's path to see his route...
```{r plot path}
cbind(move=1:length(floor),floor) %>% 
  as.data.frame() %>% 
  ggplot(aes(x=move,y=floor)) + 
  geom_step()
```

Someone's been pulling his leg! How many times does he actually pass through his final destination floor of 138 before the end of his instructions?
```{r}
which(floor==138) - 1
```

